---
title: "Propensity Score Weighting and Generalized Boosted Regression in R"
author: "Shenyang Guo and Peter Sun"
date: "September, 2021"
output: word_document
---

# Load Packages

```{r message=F, warning=F, error=F}
library(tidyverse)
library(haven)
library(sjlabelled)
library(lmtest)
library(sandwich)
library(gbm)
select <- dplyr::select
knitr::opts_chunk$set(dpi = 300,fig.width = 7)
```

# Propensity Score Weighting

## Load Data

```{r}
d <- read_dta("data/chpt5_2_original.dta") %>%
  haven::zap_formats() %>%
  sjlabelled::remove_all_labels() %>%
  as_tibble()
```

## Create ATE and ATT Weights

```{r message=FALSE, warning=FALSE}
d.weights <- d %>%
  mutate(ate_w = ifelse(kuse == 0, 1/(1-ps), 1/ps),
         att_w = ifelse(kuse == 0, ps/(1-ps), 1))
```

## PSW With ATE Weights

Use the `weights` argument in `lm()` to run OLS regression using weights and `lmtest::coeftest()` to control for clustering effects.

```{r message=FALSE, warning=FALSE}
m3 <- lm(lwss97 ~ kuse + male + black + age97 + pcged97 + mratio96, 
         data = d.weights, weights = ate_w) 
lmtest::coeftest(m3, vcov. = vcovCL(m3, cluster = d.weights$pcg_id))
```

## PSW With ATT Weights

```{r message=FALSE, warning=FALSE}
m4 <- lm(lwss97 ~ kuse + male + black + age97 + pcged97 + mratio96, 
         data = d.weights, weights = att_w)
lmtest::coeftest(m4, vcov. = vcovCL(m4, cluster = d.weights$pcg_id))
```

## Sample Imbalance Check Code

Use logistic regression for dummy covariates and OLS regression for continuous covariates. The full code can be found in Section 7.3.1 of the PSA-R code.

```{r eval=F}
c1 <- glm(male ~ kuse, family = quasibinomial, 
          data = d.weights, weights = ate_w)
lmtest::coeftest(c1, vcov. = vcovCL(c1, cluster = d.weights$pcg_id))
```

# Estimate Propensity Scores Using Generalized Boosted Regression

## Load Data and Sort

```{r}
set.seed(1000)
d2 <- read_dta("data/g3aca1.dta") %>%
  haven::zap_formats() %>%
  sjlabelled::remove_all_labels() %>%
  as_tibble() %>%
  select(intbl, ageyc, fmale, blck, whit, hisp, pcedu, ipovl, pcemft, fthr,
         dicsagg2, dicsint2, dccereg2, dccscom2, dccpros2, draggr2) %>%
  drop_na() %>%
  add_column(runif = runif(nrow(.))) %>%
  arrange(runif)
```

## Generate Propensity Scores

```{r message=F}
f5 <- as.formula(paste("intbl ~ ", paste(names(select(d2, -intbl)), 
                                         collapse = " + ")))
set.seed(1000)
m5 <- gbm::gbm(formula = f5,
          data = d2,
          distribution = "bernoulli",
          n.trees = 1000,
          train.fraction = 0.8,
          interaction.depth = 4,
          shrinkage = 0.0005)
psb <- gbm::predict.gbm(m5, data = d2, type = "response") # Create ps
```

## Plot and Summarize Results

```{r}
d2 %>%
  mutate(psb = psb,
    intbl = factor(intbl, labels = c("Control", "Treatment"))) %>%
  ggplot(aes(x = psb, color = intbl)) + theme_classic() +
  geom_density(size = 1) + xlim(0, 1) + ylim(0, 22) +
  labs(x = "Predicted Probability", y = "Density",
       title = "Propensity Scores Using Generalized Boosted Regression", color = "Treatment") +
  theme(legend.position = c(0.9, 0.9))
summary(psb)
```